Mapping Mixtures of Genotype-by-Weather Interactions in Switchgrass (Panicum virgatum)

### 2024-03-09 manuscript/revision2/revision2.tex
manuscript: New Fig 3 and Fig 4 and revised text to reflect figure changes

\begin{abstract}
The timing of vegetative and reproductive growth in plants
(``phenological timings'') depends on both genetic variation and
environmental cues; thus, phenological traits can have important
genotype-by-environment (GxE) interactions. Identifying the mixture and
prevalence of GxE primed by weather prior to the phenological event
(GxWeather) should aid prediction and manipulation of phenological
timings. Here, we map GxWeather effects on phenological timings in two
highly divergent switchgrass (\emph{Panicum virgatum}) populations using
repeated plantings of cloned individuals from these populations at eight
sites spanning the central United States. We distinguish GxWeather
patterns covarying with interpretable weather-based cues from agnostic,
site-based patterns. Most GxWeather effects belong to the latter
category and do not covary with weather-based cues. However, 65\% of
effects on the timing of vegetative growth in the Gulf population covary
with daylength 14 days prior to green-up date. 33\% of effects on the
timing of flowering in the Gulf population covary with cumulative
rainfall in the seven days prior to flowering, while 22\% of effects on
flowering timing in the Midwest population covary with day length change
two days prior to flowering. An independent pseudo-F2 cross of Gulf and
Midwest individuals at the same sites mapped 10 additive QTLs for
flowering and three for vegetative growth timing, all with significant
associations and three with enrichment of highly significant
associations in our model of GxWeather effects. We demonstrate that we
can identify genetic variation with GxWeather and assign these loci to
specific weather-based cues or other patterns. Breeding for particular
alleles at these loci could change flowering responsiveness to
photoperiod and rainfall cues in switchgrass. More broadly, this
approach could be used to identify genetic marker-environment
interactions in any species with related populations phenotyped in
multiple environments.
\end{abstract}

@@

We next looked for evidence of GxWeather by jointly re-estimating
genetic effects across all eight common gardens using the mash model,
which allows effects to covary between gardens in many ways
(Figure~\ref{fig-map} D, SI Appendix, Datasets 1-6). We specified three
categories of covariance structures across gardens: ``canonical''
covariance, with simple patterns of effect size covariance introduced in
the initial \emph{mash} manuscript; ``data-driven'' covariances derived
from common patterns of SNP effects observed in the data, and
``GxWeather'' covariance, estimated from the covariance of empirical
weather patterns at each garden at specific times before the
phenological event (Table 1, SI Appendix, Section S2). For example, the
more similar that the amount of rainfall prior to flowering is for
genotypes A and B between gardens in MI and MO, the nearer these
covariances will be to one. If rainfalls are uncorrelated, covariance
will be near zero; if there is rank changing (e.g.~high rainfall before
flowering of A at MO but not MI, and high rainfall before flowering for
B at MI but not MO), this covariance will be near negative one. We
specified multiple covariance structures for the canonical, data-driven,
and GxWeather categories, each of which represented a specific pattern
of covariance in genetic effect size estimates that could be included in
the model.

@@

The phenotypic correlations for the timing of vegetative growth had
moderate negative correlations between the Texas and North gardens,
particularly in the Gulf subpopulation and Both subpopulations
(Figure~\ref{fig-map} B). If these phenotypic correlations have a
genetic component, then they may be controlled by SNP effects that
differ in sign across these regions (effects with antagonistic
pleiotropy), SNP effects that are large in one garden or region and
non-significant in others (effects with amplification), or a combination
of these patterns. Five GxWeather covariance structures were selected by
the greedy algorithm, one to two per subpopulation. Of these five, three
had mass on them in mash models of the strong effects
(Figure~\ref{fig-covar} A-B). Two of these three matrices had negative
covariances between sites between Texas and North gardens
(Figure~\ref{fig-covar} A), while one had all positive or near-zero
covariances. SNP-associated phenotypic effects covaried with different
weather-based cues in the Gulf \& in Both subpopulations
(Figure~\ref{fig-covar} B). In total, 65\% of the posterior weight of
strong SNP effects in the mash model of Gulf vegetative growth fell on a
covariance matrix constructed using the covariance of daylength 14 days
prior to the date of vegetative growth. The covariance matrix for this
weather cue was very similar to the pattern of phenotypic correlation
for the timing of vegetative growth in the Gulf subpopulation
(Figure~\ref{fig-map} B; Figure~\ref{fig-covar} A). Mash models of the
timing of vegetative growth in the Midwest subpopulation and Both
subpopulations did not include this GxWeather covariance; the Midwest
had no weight on any GxWeather covariance, while Both subpopulations had
non-zero weights on two additional GxWeather covariance types, average
temperature one day prior to green-up, and the day length change in
seconds in the day prior to green-up (Figure~\ref{fig-covar} B). The
average temperature covariance matrix had negative covariances between
Texas and North gardens, though not as strong as the negative phenotypic
correlations seen in Both subpopulations (Figure~\ref{fig-map} B;
Figure~\ref{fig-covar} A). Only the Gulf subpopulation and Both
subpopulations had mass on any GxWeather covariance matrices
(Figure~\ref{fig-covar} C).

@@

We next characterized the pairwise patterns of effects where we were
confident in the sign of the effect at both gardens. We used the local
false sign rate (lfsr), an analogue of the lfdr that establishes
confidence in the effect sign, not the effect's difference from zero, to
determine significance. We required lfsr significance (p \textless{}
0.05) in both gardens to include effects. This means that our tests for
antagonistic pleiotropy, or a sign change between conditions, carry an
equal statistical burden to those for effects with the same sign. We
also summarized the pairwise patterns of effect sign and magnitude
within and between the Texas and North regions, within the natural
ranges of the Gulf and Midwest subpopulations
(Figure~\ref{fig-effects}).

For greenup date for the Gulf subpopulation, hundreds to thousands of
pairwise effects exhibited antagonistic pleiotropy, or a difference in
effect sign, between pairs of Texas and North gardens
(Figure~\ref{fig-effects} A). 78.7\% of pairwise comparisons between
North and Texas gardens had a difference in sign, while only 28.6\% and
0.2\% of North-North or Texas-Texas comparisons had a difference in
sign, respectively (Figure~\ref{fig-effects}). The majority of pairwise
effects for greenup for the Midwest (\textgreater55\%) and Both
(\textgreater85\%) subpopulations were the same sign, and effects
frequently differed in magnitude between the MO and OK garden and other
gardens (Figure~\ref{fig-effects} X).

For flowering date for the Gulf subpopulation, less than 2\% of pairwise
effects exhibited antagonistic pleiotropy, or a difference in effect
sign, within or between regions (Figure~\ref{fig-effects}). More effects
differed in magnitude between the Texas and North regions than within
these regions (42.7\% vs \textless20\%; Figure~\ref{fig-effects} X). The
Midwest population had relatively few significant effects for flowering,
but a large proportion of these differed in sign between Texas and North
regions (42.7\%) or within the North region (65.4\%). Finally, in Both
subpopulations, less than 20\% of pairwise effects differed in sign
(Figure~\ref{fig-effects}). Most differences in sign were between TX1,
the southernmost garden, and all other gardens (Figure~\ref{fig-effects}
B). Similarly, more effects that differed in magnitude included gardens
in the Texas region (52.3-55.9\%), and most effect pairs in the North
region were not distinguishable (91.5\%).

@@

We sought additional experimental support for our joint re-estimates of
SNP effects using an independent pseudo-F2 mapping population created
from Gulf \& Midwest individuals and grown at the same sites
(Figure~\ref{fig-qtl} A,B). We conducted quantitative trait loci (QTL)
mapping of flowering as functions of five environmental cues that we
also used as covariance matrices in mash, and identified eight QTL for
flowering date, eight QTL for flowering as a function of day length
change two days prior, one QTL for the start of vegetative growth, and
two QTL for vegetative growth as a function of daylength change one day
prior, all of which showed QTL by environment interactions (SI Appendix,
Fig. S3). All QTL for flowering overlapped one or more homologs from
rice or \emph{A. thaliana} with functionally validated roles in
flowering (SI Appendix, Dataset 7). All flowering and green-up QTL
intervals contained at least one SNP significant in at least one mash
run at a log10-transformed Bayes Factor \textgreater{} 2, or in the 1\%
tail of significance, whichever was stricter (SI Appendix, Dataset 8).
We also looked for enrichments of mash SNPs in the 1\% tail of
significance (the `mash 1\% tail') within each QTL interval. At the 5\%
level, three QTL had enrichments of SNPs in the mash 1\% tail. Overall,
there were five significant enrichments (p \textless{} 0.05,
hypergeometric test) of SNPs in the mash 1\% tail in the QTL intervals;
however, this number of enrichments was found in many random genomic
intervals also (Figure~\ref{fig-qtl} C). Thus, we were able to
experimentally support some of our re-estimates of SNP effects with a
QTL mapping experiment using a separate mapping population.

@@

Our analysis of the timing of vegetative growth in the Gulf and in Both
subpopulations revealed substantial antagonistic pleiotropy in effects
between the Texas and North gardens (@fig-effects A). This result
supports theoretical models that local adaptation should involve
antagonistic pleiotropy at the level of individual loci (22--25), and is
the first experimental work using GWAS across common gardens to find
antagonistic pleiotropy to be common in small genomic regions (12, 37,
38).

Our analysis of the timing of flowering showed that the Gulf and Midwest
subpopulations have distinct GxWeather: flowering timing in the Midwest
subpopulation has photoperiod-related variation, in that flowering
timing covaries with a day length change signal two days before
flowering occurs. In contrast, the Gulf subpopulation does not have
variation in flowering based on a photoperiod cue. Instead, the Gulf
subpopulation has variation in flowering that covaries with the rainfall
that occurs in the week prior to flowering. Three genomic regions
affecting flowering that we re-estimated across all eight sites were
also supported by QTL from an independent mapping population at these
sites (@fig-qtl A).

Identifying the environmental cues that are predictive of, or even
correlated with, plant phenotypic responses remains a major challenge to
studies interrogating gene action across many natural environments. The
GxWeather photoperiod and cumulative rainfall cues we identify here are
functions of the genotypes measured and capture only a minority of SNP
effects on flowering. We know still less about the overwintering
parameters that cause variation in the start of vegetative growth.
Though we could assign SNP effects to a GxWeather covariance structure
in the Gulf subpopulation, we could not do this for the Midwest
subpopulation. More generally, it is difficult to predict the time
scales over which individuals may integrate environmental cues,
particularly in perennial species which may integrate these cues over
longer time scales. If this integration time itself varies between
individuals, we cannot select a covariance structure that reflects this,
though these structures would likely be highly correlated with GxWeather
structures we did include. Mash offers an opportunity to specify
multiple environmental cues and compete them to explain patterns of
genetic effects, allowing us to detect how important these cues are
genome-wide, and how strongly each cue influences each SNP. This is a
key development to further improve our understanding of genetic
variation in GxE.

@@

\includegraphics{images/Figure_3_Effect_Types_by_Site.png}

@@

\caption{\label{fig-effects}AHM is considering moving this figure to the
supplement and substantially simplifying the figure in the main text,
perhaps just showing total effects in Texas, the North, and maybe
Oklahoma. Comparisons of pairs of jointly re-estimated effects across
eight gardens, for effects with lfsr \textless{} 0.05. Essentially
making Table 2 into a new, simplified figure. Effect patterns for
re-estimated SNP effects that are significant at pairs of sites by the
local false sign rate test (lfsr), for (a) greenup date and (b)
flowering date, within and between two genetic subpopulations. Common
gardens are arranged in latitudinal order along the x- and y-axis.
Effects in the first and fourth rows differ in sign at these pairs of
gardens (p \textless{} 0.05, lfsr). Effects in the second, third, fifth
and sixth rows are identical in sign (p \textless{} 0.05, lfsr). Effects
in the second and fifth rows additionally differ in magnitude by a
factor of \textgreater0.4, while effects in the third and sixth row were
not distinguishable by magnitude nor sign of the effect.}

@@

\includegraphics{images/Figure_4_mash_SNP_effects.png}

@@

\caption{\label{fig-qtl}SNP effects for SNPs within QTL. AHM to re-make
this figure now LZ has re-run the QTL analysis, the assmption being the
re-analysis will not change the figure or the text much. The bigger
question for TEJ and AH is whether we want to keep this analysis in the
paper, what does it add?}

@@

To evaluate the prevalence and kinds of covariance patterns of SNP
effects across our common gardens, we used multivariate adaptive
shrinkage (\emph{mash}) on SNP effect estimates from the diversity panel
(35). To do this, we first conducted univarate genome-wide association
(GWAS) on site-specific best linear unbiased predictors (BLUPs) for the
timing of vegetative growth and flowering, using the switchgrassGWAS
package and the methods in (39). We used the effect estimates for single
nucleotide polymorphisms (SNPs) from these GWAS as preliminary effect
estimates for both traits at each garden. We then jointly modeled
genetic effect estimates, and the mixture of ways these effects might
covary, across all eight common gardens, using a two-part procedure. We
first modeled genetic effects with different combinations of covariance
structures, using the effect size estimates from a subset of \ul{19K}
relatively unlinked (r\textsuperscript{2} \textless{} 0.2), randomly
selected SNPs used in the GWAS. In these models (hereafter referred to
as ``random''), we then specified a variety of potential covariance
structures among genetic effect size estimates from each garden (as
expanded on below) and used a greedy algorithm to iteratively select and
add covariance structures that significantly improved the log likelihood
of the \emph{mash} model. Given the high correlation between some of the
covariance patterns we tested, this approach allowed us to select only
the subset of patterns of covariance that significantly improved the fit
of the jointly inferred structure of genetic effect estimates present in
the data. We then selected 19K relatively unlinked (r\textsuperscript{2}
\textless{} 0.2) SNPs from each univariate GWAS that had the largest
effect estimates in each GWAS (hereafter ``strong'' effects), which gave
a set of \ul{\textbf{48K - 80K}} SNPs, and jointly re-estimated these
effect size estimates at all eight sites using the set of significant
covariance matrices that most improved the model fit on the random
effects. Notably, the majority of these ``strong'' effects were not
significant in univariate GWAS.

@@

In addition, the get\_GxE function also sets an arbitrary threshold to
count an effect as changing in magnitude between environments, commonly
known as differential sensitivity or a change in amplitude of . For
differential sensitivity, this function determines if effects
significant in both conditions are of the same sign and of a magnitude
(not tested for significance) that differs by a factor of 0.4 or more.
The remaining effects that are significant in both conditions have the
same effect sign and similar effect magnitudes and we denote these
effects as having no GxE. The distinction between effects with different
magnitudes is arbitrary but useful to fully characterize how effects
vary across environments. and our use of the lfsr to determine
significance and our specification that SNP effects must be significant
in both conditions to be included means that our tests for antagonistic
pleiotropy carry an equal statistical burden to those measuring
differential sensitivity and effects without GxE.

@@

To confirm candidate genomic regions and patterns of allelic effects
found in the diversity panel, we analyzed flowering in an outbred
pseudo-F2 cross between four individuals, two Midwest and two Gulf
individuals. The formation of this mapping population has been described
previously (34); additional details on QTL mapping can be found in SI
Appendix, Section S6. To be directly comparable to the diversity panel
data, only 2019 phenology data from the pseudo-F2 cross from the same
eight common garden sites were used. To compare our QTL enrichments of
significant mash associations to the null expectation, we used
permutation to choose 1000 sets of 23 genomic regions of the same size
randomly distributed throughout the genome, then calculated enrichments
of the mash 1\% tail in these random intervals.

>>>>

\begin{abstract}
The timing of vegetative and reproductive growth in plants
(``phenological timings'') depends on both genetic variation and
environmental cues; thus, phenological traits can have important
genotype-by-environment (GxE) interactions. Identifying the mixture and
prevalence of GxE primed by weather prior to the phenological event
(GxWeather) should aid prediction and manipulation of phenological
timings. Here, we map GxWeather effects on phenological timings in two
highly divergent switchgrass (\emph{Panicum virgatum}) populations using
repeated plantings of cloned individuals from these populations at eight
sites spanning the central United States. We distinguish GxWeather
patterns covarying with interpretable weather-based cues from agnostic,
site-based patterns. Most GxWeather effects belong to the latter
category and do not covary with weather-based cues. However, 65\% of
effects on the timing of vegetative growth in the Gulf population covary
with daylength 14 days prior to green-up date. 33\% of effects on the
timing of flowering in the Gulf population covary with cumulative
rainfall in the seven days prior to flowering, while 22\% of effects on
flowering timing in the Midwest population covary with day length change
two days prior to flowering. An independent pseudo-F2 cross of Gulf and
Midwest individuals at the same sites mapped eight additive QTLs for
flowering and three for vegetative growth timing, all with significant
associations and three with enrichment of highly significant
associations in our model of GxWeather effects. We demonstrate that we
can identify genetic variation with GxWeather and assign these loci to
specific weather-based cues or other patterns. Breeding for particular
alleles at these loci could change flowering responsiveness to
photoperiod and rainfall cues in switchgrass. More broadly, this
approach could be used to identify genetic marker-environment
interactions in any species with related populations phenotyped in
multiple environments.
\end{abstract}

@@

We next looked for evidence of GxWeather by jointly re-estimating
genetic effects across all eight common gardens using the mash model,
which allows effects to covary between gardens in many ways
(Figure~\ref{fig-map} D, SI Appendix, Datasets 1-6). We specified three
categories of covariance structures across gardens: ``canonical''
covariance, with simple patterns of effect size covariance introduced in
the initial \emph{mash} manuscript; ``data-driven'' covariances derived
from common patterns of SNP effects observed in the data, and
``GxWeather'' covariance, estimated from the covariance of empirical
weather patterns at each garden at specific times before the
phenological event (Table 1, SI Appendix, Section S2). For example, the
more similar that the amount of rainfall prior to flowering is for
genotypes A and B between gardens in MI and MO, the nearer these
covariances will be to one. If rainfalls are uncorrelated, covariance
will be near zero; if there is rank changing (e.g.~high rainfall before
flowering of A at MO but not MI, and high rainfall before flowering for
B at MI but not MO), this covariance will be near negative one. We
specified multiple covariance structures for the canonical, data-driven,
and GxWeather categories, each of which represented a specific pattern
of covariance in genetic effect size estimates that could be included in
the model. The phenotypic correlations for the timing of vegetative
growth had moderate negative correlations between the Texas and North
gardens, particularly in the Gulf subpopulation and Both subpopulations
(Figure~\ref{fig-map} B). If these phenotypic correlations have a
genetic component, they could be partially or completely captured by the
covariance structures specified in the mash model.

@@

Five GxWeather covariance structures were selected by the greedy
algorithm, zero to two per subpopulation: two for the start of
vegetative growth, and three for flowering date. Of these five, three
had mass on them in mash models of the strong effects
(Figure~\ref{fig-covar} A-B). Two of these three matrices had negative
covariances between sites between Texas and North gardens
(Figure~\ref{fig-covar} A), while one had all positive or near-zero
covariances. SNP-associated phenotypic effects covaried with different
weather-based cues in the Gulf \& in Both subpopulations
(Figure~\ref{fig-covar} B). In total, 65\% of the posterior weight of
strong SNP effects in the mash model of Gulf vegetative growth fell on a
covariance matrix constructed using the covariance of daylength 14 days
prior to the date of vegetative growth. The covariance matrix for this
weather cue was very similar to the pattern of phenotypic correlation
for the timing of vegetative growth in the Gulf subpopulation
(Figure~\ref{fig-map} B; Figure~\ref{fig-covar} A). Mash models of the
timing of vegetative growth in the Midwest subpopulation and Both
subpopulations did not include this GxWeather covariance; the Midwest
had no weight on any GxWeather covariance, while Both subpopulations had
non-zero weights on two additional GxWeather covariance types, average
temperature one day prior to green-up, and the day length change in
seconds in the day prior to green-up (Figure~\ref{fig-covar} B). The
average temperature covariance matrix had negative covariances between
Texas and North gardens, though not as strong as the negative phenotypic
correlations seen in Both subpopulations (Figure~\ref{fig-map} B;
Figure~\ref{fig-covar} A). Only the Gulf subpopulation and Both
subpopulations had mass on any GxWeather covariance matrices
(Figure~\ref{fig-covar} C).

@@

We next characterized the pairwise patterns of effects where we were
confident in the sign of the effect at both gardens. If the overall
pattern of phenotypic expression differs between a pair of gardens, this
difference may be comprised of SNP effects that differ in sign between
these gardens (effects with antagonistic pleiotropy), SNP effects that
differ in magnitude - that are large in one garden or region and smaller
in others, and SNP effects that are indistinguishable in the two gardens
(Figure~\ref{fig-effects} A). We used the local false sign rate (lfsr),
an analogue of the lfdr that establishes confidence in the effect sign,
not the effect's difference from zero, to determine significance. We
required lfsr significance (p \textless{} 0.05) in both gardens to
include effects. This means that our tests for antagonistic pleiotropy,
or a sign change between conditions, carry an equal statistical burden
to those for effects with the same sign.

For greenup date for the Gulf subpopulation, hundreds to thousands of
pairwise effects exhibited antagonistic pleiotropy, or a difference in
effect sign, between pairs of Texas and North gardens
(Figure~\ref{fig-effects} A; (\textbf{SI-fig-gardens?})). 78.7\% of
pairwise comparisons between North and Texas gardens had a difference in
sign, while only 28.6\% and 0.2\% of North-North or Texas-Texas
comparisons had a difference in sign, respectively
(Figure~\ref{fig-effects}). The majority of pairwise effects for greenup
for the Midwest (\textgreater55\%) and Both (\textgreater85\%)
subpopulations were the same sign, and effects most often differed in
magnitude between gardens within and between the regions
(Figure~\ref{fig-effects} X).

For flowering date for the Gulf subpopulation, less than 2\% of pairwise
effects exhibited antagonistic pleiotropy, or a difference in effect
sign, within or between regions (Figure~\ref{fig-effects}). More effects
differed in magnitude between the Texas and North regions than within
these regions (42.7\% vs \textless20\%; Figure~\ref{fig-effects} X),
while the majority of effects had no GxE. The Midwest population had
relatively few significant effects for flowering, but a large proportion
of these differed in sign between Texas and North regions (42.7\%) or
within the North region (65.4\%). Finally, in Both subpopulations, less
than 20\% of pairwise effects differed in sign
(Figure~\ref{fig-effects}). Most differences in sign were between TX1,
the southernmost garden, and all other gardens
((\textbf{SI-fig-gardens?})). Similarly, more effects that differed in
magnitude included gardens in the Texas region (52.3-55.9\%), and most
effect pairs in the North region were not distinguishable (91.5\%).

@@

We sought additional experimental support for our joint re-estimates of
SNP effects using an independent pseudo-F2 mapping population created
from Gulf \& Midwest individuals and grown at the same sites
(Figure~\ref{fig-qtl} A,B). We conducted quantitative trait loci (QTL)
mapping of flowering as functions of five environmental cues that we
also used as covariance matrices in mash, and identified eight QTL for
flowering date, eight QTL for flowering as a function of day length
change two days prior, one QTL for the start of vegetative growth, and
two QTL for vegetative growth as a function of daylength change one day
prior, all of which showed QTL by environment interactions (SI Appendix,
Fig. S3). All QTL for flowering overlapped one or more homologs from
rice or \emph{A. thaliana} with functionally validated roles in
flowering (SI Appendix, Dataset 7). All flowering and green-up QTL
intervals contained at least one SNP significant in at least one mash
run at a log10-transformed Bayes Factor \textgreater{} 2, or in the 1\%
tail of significance, whichever was stricter (SI Appendix, Dataset 8).
We also looked for enrichments of mash SNPs in the 1\% tail of
significance (the `mash 1\% tail') within each QTL interval. At the 5\%
level, three QTL had enrichments of SNPs in the mash 1\% tail. Overall,
there were five significant enrichments (p \textless{} 0.05,
hypergeometric test) of SNPs in the mash 1\% tail in the QTL intervals.
Thus, we were able to experimentally support some of our re-estimates of
SNP effects with a QTL mapping experiment using a separate mapping
population.

@@

Our analysis of the timing of vegetative growth in the Gulf and in Both
subpopulations revealed substantial antagonistic pleiotropy in effects
between the Texas and North gardens (Figure~\ref{fig-effects} B). This
result supports theoretical models that local adaptation should involve
antagonistic pleiotropy at the level of individual loci (22--25), and is
the first experimental work using GWAS across common gardens to find
antagonistic pleiotropy to be common in small genomic regions (12, 37,
38).

Our analysis of the timing of flowering showed that the Gulf and Midwest
subpopulations have distinct GxWeather: flowering timing in the Midwest
subpopulation has photoperiod-related variation, in that flowering
timing covaries with a day length change signal two days before
flowering occurs. In contrast, the Gulf subpopulation does not have
variation in flowering based on a photoperiod cue. Instead, the Gulf
subpopulation has variation in flowering that covaries with the rainfall
that occurs in the week prior to flowering. Three genomic regions
affecting flowering that we re-estimated across all eight sites were
also supported by QTL from an independent mapping population at these
sites (Figure~\ref{fig-qtl} B).

Identifying the environmental cues that are predictive of, or even
correlated with, plant phenotypic responses remains a major challenge to
studies interrogating gene action across many natural environments. The
GxWeather photoperiod and cumulative rainfall cues we identify here are
functions of the genotypes measured and capture only a minority of SNP
effects on flowering. We know still less about the overwintering
parameters that cause variation in the start of vegetative growth. We
could only assign SNP effects to a GxWeather covariance structures in
four of the six phenotype \& genetic subpopulations we modeled. More
generally, it is difficult to predict the time scales over which
individuals may integrate environmental cues, particularly in perennial
species which may integrate these cues over longer time scales. If this
integration time itself varies between individuals, we cannot select a
covariance structure that reflects this, though these structures would
likely be highly correlated with GxWeather structures we did include.
Mash offers an opportunity to specify multiple environmental cues and
compete them to explain patterns of genetic effects, allowing us to
detect how important these cues are genome-wide, and how strongly each
cue influences each SNP. This is a key development to further improve
our understanding of genetic variation in GxE.

@@

\includegraphics{Figure_3_GxE.png}

@@

\caption{\label{fig-effects}Types of GxE present between of pairs of
jointly re-estimated effects in eight common gardens, for effects with
lfsr \textless{} 0.05. a) Examples of effect patterns at three pairs of
sites with three types of GxE. Sign: Effects that differ in sign at
these pairs of gardens (p \textless{} 0.05, lfsr). Magnitude: Effects
identical in sign (p \textless{} 0.05, lfsr) that differ in magnitude by
a factor of \textgreater0.4. None: Effects not distinguishable by
magnitude nor sign of the effect. b) The fraction of effects with each
GxE type for the start of vegetative growth (Greenup Date) and
reproductive growth (Flowering Date), within and between two genetic
subpopulations. Common gardens are grouped by the larger region they
came from: North gardens are within the natural range of the Midwest
subpopulation, while Texas gardens are within the natural range of the
Gulf subpopulation.}

@@

\includegraphics{Figure_4_QTL_Overlaps.png}

@@

\caption{\label{fig-qtl}Overlaps of QTL from an outbred pseudo-F2 cross
and with jointly re-estimated SNP effects in the 1\% tail of
significance from a diversity panel. Dotted lines indicate
permutation-based significance thresholds for each weather-related
function. Stars indicate QTL with significant enrichment for SNPs in the
1\% mash tail; G, M, and B indicate which subpopulation had enrichment:
G - Gulf subpopulation, M -Midwest subpopulation, B - both
subpopulations. Rug plots show genomic locations of SNPs in the 1\% mash
tail for flowering date for each subpopulation. a) QTL mapping for the
start of vegetative growth (Greenup Date), and three weather-related
functions of greenup date. b) QTL mapping for the start of reproductive
growth (Flowering Date), and two weather-related functions of flowering
date.}

@@

To evaluate the prevalence and kinds of covariance patterns of SNP
effects across our common gardens, we used multivariate adaptive
shrinkage (\emph{mash}) on SNP effect estimates from the diversity panel
(35). To do this, we first conducted univarate genome-wide association
(GWAS) on site-specific best linear unbiased predictors (BLUPs) for the
timing of vegetative growth and flowering, using the switchgrassGWAS
package and the methods in (39). We used the effect estimates for single
nucleotide polymorphisms (SNPs) from these GWAS as preliminary effect
estimates for both traits at each garden. We then jointly modeled
genetic effect estimates, and the mixture of ways these effects might
covary, across all eight common gardens, using a two-part procedure. We
first modeled genetic effects with different combinations of covariance
structures, using the effect size estimates from a subset of \ul{19K}
relatively unlinked (r\textsuperscript{2} \textless{} 0.2), randomly
selected SNPs used in the GWAS. In these models (hereafter referred to
as ``random''), we then specified a variety of potential covariance
structures among genetic effect size estimates from each garden (as
expanded on below) and used a greedy algorithm to iteratively select and
add covariance structures that significantly improved the log likelihood
of the \emph{mash} model. Given the high correlation between some of the
covariance patterns we tested, this approach allowed us to select only
the subset of patterns of covariance that significantly improved the fit
of the jointly inferred structure of genetic effect estimates present in
the data. We then selected 19K relatively unlinked (r\textsuperscript{2}
\textless{} 0.2) SNPs from each univariate GWAS that had the largest
effect estimates in each GWAS (hereafter ``strong'' effects), which gave
sets of \ul{\textbf{48K - 80K}} SNPs, and jointly re-estimated these
effect size estimates at all eight sites using the set of significant
covariance matrices that most improved the model fit on the random
effects. Notably, the majority of these ``strong'' effects were not
significant in univariate GWAS.

@@

In addition, the get\_GxE function also sets an arbitrary threshold to
count an effect as changing in magnitude between environments, commonly
known as differential sensitivity or a change in amplitude of the
effect. For differential sensitivity, this function determines if
effects significant in both conditions are of the same sign and of a
magnitude (not tested for significance) that differs by a factor of 0.4
or more. The remaining effects that are significant in both conditions
have the same effect sign and similar effect magnitudes and we denote
these effects as having no GxE. The distinction between effects with
different magnitudes is arbitrary but useful to fully characterize how
effects vary across environments to ultimately influence phenotypes. Our
use of the lfsr to determine significance and our specification that SNP
effects must be significant in both conditions to be included means that
our tests for antagonistic pleiotropy carry an equal statistical burden
to those measuring differential sensitivity and effects without GxE.

@@

To confirm candidate genomic regions and patterns of allelic effects
found in the diversity panel, we analyzed flowering in an outbred
pseudo-F2 cross between four individuals, two Midwest and two Gulf
individuals. The formation of this mapping population has been described
previously (34); additional details on QTL mapping can be found in SI
Appendix, Section S6. To be directly comparable to the diversity panel
data, only 2019 phenology data from the pseudo-F2 cross from the same
eight common garden sites were used.
